---
title: "p8105_hw1_lc4152"
author: "Lin Cao"
date: "2026-09-21"
output: github_document
---

# Problem 1

```{r}
library(tidyverse)
```

```{r}
data("penguins", package = "palmerpenguins")
penguins
```
The `penguins` dataset contains information about penguins, including their species (`r paste(unique(penguins$species), collapse = ', ')`); island (`r paste(unique(penguins$island), collapse = ', ')`); bill length; bill depth; flipper length; body mass; sex (`r paste(unique(penguins$sex), collapse = ', ')`); and year of observation (`r paste(unique(penguins$year), collapse = ', ')`).

The size of the dataset is `r nrow(penguins)` observations and `r ncol(penguins)` variables. The mean flipper length is `r round(mean(penguins$flipper_length_mm, na.rm = TRUE), 2)` mm.

```{r}
penguins_plot = 
  ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point()
penguins_plot
```

```{r}
ggsave("scatterplot.png", plot = penguins_plot)
```


## Problem 2
```{r}
set.seed(1)
problem2_df = 
  tibble(
    sample_random = rnorm(10),
    vec_positive = sample_random > 0,
    vec_char = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j"),
    vec_factor = factor(c("A", "B", "A", "A", "C", "B", "C", "B", "B", "A"))
  )
problem2_df
```

Take the mean of each variable.

```{r}
mean(pull(problem2_df, sample_random))
mean(pull(problem2_df, vec_positive))
mean(pull(problem2_df, vec_char))
mean(pull(problem2_df, vec_factor))
```

The mean works for `sample_random` and `vec_positive`, which are numeric and logical variables. It doesn't work for `vec_char` and `vec_factor`.

Apply `as.numeric` to logical, char, and factor.

```{r, results='hide', warning=FALSE}
as.numeric(pull(problem2_df, vec_positive))
as.numeric(pull(problem2_df, vec_char))
as.numeric(pull(problem2_df, vec_factor))
```


Converting the logical variable to numeric changes `TRUE` to 1 and `FALSE` to 0. This explains why taking the mean of a logical variable works.

For character, it produces `NA` values because the character strings cannot be interpreted as numbers.

For factor variable, it returns the integer codes corresponding to the factor levels. 

Therefore, these numbers represent the internal coding of the factor levels and do not necessarily have a meaningful numerical interpretation.
















